\documentclass{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage{proof}
\usepackage{txfonts}
\usepackage{mathtools}

\newtheorem{thm}{Theorem}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{defn}[thm]{Definition}

\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

\def\imp{\to}
\def\land{\mathbin\&}


\title{Automated Deduction in Matching Logic}
\author{FSL group}



\begin{document}
\maketitle

Recent success in building very fast automated theorem provers (especially for first-order theories) makes us look forward to an highly efficient automated deductive system for matching logic. This project aims at that.

\section{Grammar}

As in most logic, formulas of matching logic, called \emph{patterns}, are written in a formal language, denoted as $\mathcal{L}$, who has a very similar grammar as first order logic. 

The language $\mathcal{L}$ in general is a many-sorted language. A signature of $\mathcal{L}$ contains not only a finite set $\Sigma$ of symbols, but also a finite nonempyty set $S$ of sorts. Each symbol $\sigma \in \Sigma$ is, of course, sorted, with a fixed nonempty arity. We write $\sigma \in \Sigma_{s_1,\dots,s_n,s}$ when we want to emphasize that $\sigma$ takes $n$ arguments (with suggested sorts) and returns a pattern in sort $s$, but we hope in most cases sorting is clear from context.

The basic grammar for $\mathcal{L}$, as defined below, is almost identical to first-order logic, except that in $\mathcal{L}$ there is no difference between relational (predicate) and functional symbols, and we accept first-order terms as patterns in matching logic. 
\begin{align*}
P \Coloneqq\  & x \\
       \mid\  & P_1 \to P_2 \\
       \mid\  & \neg P \\
       \mid\  & \forall x . P \\
       \mid\  & \sigma(P_1,\dots,P_n).
\end{align*}
For simplicity, we did not mention sorting in the grammar definition, and assume it should be clear to all readers. For example, in $P_1 \to P_2$, both patterns $P_1$ and $P_2$ should have the same sort, and that sort is the sort of $P_1 \to P_2$. The sort of $\forall x . P$ is the sort of $P$, where the sort of variable $x$ does not matter. To see why it is the case, consider the pattern $\exists x . list(x, 1 \cdot 3 \cdot 5)$, which is the set of all memory configurations that has a list $(1,3,5)$ in it.

Propositional connectives are always assumed, including conjunction ($\wedge$), disjunction ($\vee$), and equivalence ($\leftrightarrow$). Existential quantifier ($\exists x$) is defined by universal quantifier ($\forall x$) in the normal way. Bottom ($\bot_s$) and top ($\top_s$) in sort $s$ are given by $x \wedge \neg x$ and $\neg \bot_s$, respectively, where $x$ is a variable in sort $s$. It does not matter which variable we pick.

\subsection{Extension}
The formal language is often extended with definedness symbols. A definedness symbol $\ceil*{\_}_{s_1}^{s_2} \in \Sigma_{s_1,s_2}$ is a \emph{predicate} symbol, i.e., $\ceil*{P}_{s_1}^{s_2}$ is either $\top$ or $\bot$ for any pattern $P$. We often write $\ceil*{P}$ when sorting information can be derived from the context.

Definedness extends the formal language:
\begin{align*}
&\floor*{P} \coloneqq {\neg} \ceil*{\neg P} \\
&P_1 = P_2 \coloneqq \floor*{P_1 \leftrightarrow P_2} \\
&P_1 \neq P_2 \coloneqq \neg (P_1 = P_2) \\
&P_1 \subseteq P_2 \coloneqq \floor*{P_1 \to P_2} \\
&x \in P \coloneqq x \subseteq P
\end{align*}
and provides convenient ways to write \emph{predicate patterns}, patterns that is either $\top$ or~$\bot$. The sort of a predicate patterns is often of no importance. It is its ``truth value'' that is important. For example, $\ceil*{P}$ is ``true'' if $P$ is not empty, and this fact can be used in any context. This leads us to polymorphic sorting in matching logic.

\subsection{Polymorphism}
Predicate symbols are often polymorphic sorting. As a result, many predicate connectives are of polymorphic types, too. Predicate pattern $P_1 = P_2$ only requires $P_1$ and $P_2$ have the same sort, and its result can be of any sort. If $P_1$ and $P_2$ have different sorts, $P_1 = P_2$ is \emph{ill-sorted}.

Polymorphism means in mathematics a mapping from $\mathcal{Patterns}$


\section{Hilbert proof system}
Axioms in $\mathcal{L}$ are given by the following nine axiom schemata where $P, Q, R$ are arbitrary patterns and $x, y$ are variables.
\begin{itemize}
\item (K1) $P \to (Q \to P)$
\item (K2) $(P \to (Q \to R)) \to ((P \to Q) \to (P \to R))$
\item (K3) $(\neg P \to \neg Q) \to (Q \to P)$
\item (K4) $\forall x . (P \to Q) \to (P \to \forall x . Q)$ if $x$ does not occur free in $P$
\item (K5) $\exists y . x = y$
\item (K6) $\exists y . Q = y \to (\forall x . P(x) \to P[Q/x])$ if $Q$ is free for $x$ in $P$
\item (K7) $P_1 = P_2 \to (Q[P_1/x] \to Q[P_2/x])$
\item (M1) $x \in y = (x = y)$
\item (M2) $x \in \neg P = \neg (x \in P)$
\item (M3) $x \in P \wedge Q = (x \in P) \wedge (x \in Q)$
\item (M4) $x \in \exists y . P = \exists y . x \in P$ where $x$ is distinct from $y$
\item (M5) $x \in \sigma(\dots,P_i,\dots) = \exists y . y \in P_i \wedge x \in \sigma(\dots,y,\dots)$
\end{itemize}

Inference rules include
\begin{itemize}
\item (Modus Ponens) From $P$ and $P \to Q$, deduce $Q$.
\item (Universal Generalization) From $P$, deduce $\forall x . P$. 
\item (Membership Introduction) From $P$, deduce $\forall x . (x \in P)$, where $x$ does not occur free in $P$.
\item (Membership Elimination) From $\forall x . (x \in P)$, deduce $P$, where $x$ does not occur free in $P$.
\end{itemize}

\begin{thm}[Soundness of $K_\mathcal{L}$]
Theorems of $K_\mathcal{L}$ are valid.
\end{thm}
\begin{proof}
	Trivial.
\end{proof}

We provide some metatheorems of $K_\mathcal{L}$.

\begin{prop}[Tautology] \label{prop:taut}
	For any propositional tautology $\mathcal{A}(p_1,\dots,p_n)$ where $p_1,\dots,p_n$ are propositional variables, $$ \vdash \mathcal{A}(P_1,\dots,P_n).$$
\end{prop}
\begin{proof}
	Omit proof here.
\end{proof}
\paragraph{Remark} Proposition~\ref{prop:taut} makes any metatheorem of propositional logic a metatheorem of $K_\mathcal{L}$.
\begin{prop} [Variable Substitution]
	$\vdash \forall x . P \to P[y / x]$.
\end{prop}
\begin{prop}[Functional Substitution]
	$\vdash \exists y . (Q = y) \to (P[Q / x] \to \exists x . P(x))$.
\end{prop}
\begin{prop}[$\vee$-Introduction] \label{prop:vee-introduction}
	$\vdash P$ implies $\vdash P \vee Q$.
\end{prop}
\begin{proof}
Use Proposition~\ref{prop:taut} and Modus Ponens.
Note that in general, $\vdash P \vee Q$ does not imply $\vdash P$ or $\vdash Q$.
\end{proof}

\begin{prop}[$\wedge$-Introduction and Elimination]
\label{prop:wedge}
$\vdash P$ and $\vdash Q$ iff $\vdash P \wedge Q$.
\end{prop}
\begin{proof}
Use Proposition~\ref{prop:taut} and Modus Ponens.
\end{proof}

\begin{prop}[Equality Introduction]
	$\vdash P = P$.
\end{prop}
\begin{proof}
	Use Membership Introduction and Proposition~\ref{prop:taut}.
\end{proof}

\begin{prop}[Equality Replacement]
\label{prop:ereplacement}
$\vdash P_1 = P_2$ and $\vdash Q[P_1 / x]$ implies $\vdash Q[P_2 / x]$.
\end{prop}
\begin{proof}
	Use Axiom~(K7) and Modus Ponens.
\end{proof}

\begin{prop}[Equality Establishment]
\label{prop:eestablish}
$\vdash P \leftrightarrow Q$ implies $\vdash P = Q$.
\end{prop}
\begin{proof}
	Use Membership Axoims and $\vee$-Introduction.
\end{proof}
\begin{cor} \label{cor:top}
$\vdash P$ implies $\vdash P = \top$ .
\end{cor}

\begin{prop} \label{prop:var1}
$\vdash x \in \ceil*{y}$.
\end{prop}
\begin{proof}
\begin{align*}
&\vdash x \in \ceil*{y} \\
{\text{if}}& \vdash \forall x . (x \in \ceil*{y}) \tag{$K5$, $K6$, and Modus Ponens} \\
{\text{iff}}& \vdash \ceil*{y}.
\end{align*}
\end{proof}

\begin{prop} \label{prop:p->ceilp}
$\vdash P \to \ceil*{P}$.
\end{prop}
\begin{proof}
\begin{align*} 
&\vdash P \to \ceil*{P} \\
{\text{iff}}& \vdash \forall x . (x \in P \to \ceil*{P}) \\
{\text{if}}&  \vdash x \in P \to \ceil*{P} \\
{\text{iff}}& \vdash x \in P \to x \in \ceil*{P} \\
{\text{iff}}& \vdash x \in P \to \exists y . (y \in P \wedge x \in \ceil*{y}) \\
{\text{iff}}& \vdash x \in P \to \neg \forall y . (y \not \in P \vee x \not \in \ceil*{y}) \\
{\text{iff}}& \vdash \forall y . (y \not \in P \vee x \not \in \ceil*{y}) \to x \not \in P \\
{\text{if}}&  \vdash x \not \in P \vee x \not \in \ceil*{x} \to x \not \in P \\
{\text{iff}}& \vdash x \in P \to x \in P \wedge x \in \ceil*{x} \\
{\text{iff}}& \vdash x \in P \to x \in \ceil*{x} \\
{\text{if}}&  \vdash x \in \ceil*{x} \\
\end{align*}
\paragraph{Remark} Similarly we can show $\vdash \floor*{P} \to P$.
\end{proof}

\begin{prop} \label{prop:floor}
$\vdash \forall x . (x \in P) = \floor*{P}$, where $x$ occurs free in $P$.
\end{prop}
\begin{proof}
By Proposition~\ref{prop:eestablish} and~\ref{prop:wedge}, it suffices to show \begin{equation} \label{eq:prop11-1}
\vdash \forall x . (x \in P) \to \floor*{P}
\end{equation}
and
\begin{equation} \label{eq:prop11-2}
\vdash \floor*{P} \to \forall x . (x \in P).
\end{equation}

To show~\eqref{eq:prop11-1},

\begin{align*}
&\vdash \forall x . (x \in P) \to \floor*{P} \\
{\text{iff}}& \vdash \forall x . \ceil*{x \wedge P} \to \neg \ceil*{\neg P} \\
{\text{iff}}& \vdash \ceil*{\neg P} \to \exists x . \neg \ceil*{x \wedge P} \\
{\text{iff}}& \vdash \forall y . (y \in ( \ceil*{\neg P} \to \exists x . \neg \ceil*{x \wedge P})) \\
{\text{if}}& \vdash y \in ( \ceil*{\neg P} \to \exists x . \neg \ceil*{x \wedge P} \\
{\text{iff}}& \vdash \exists z_1 . (z_1 \not \in P \wedge y \in \ceil*{z_1}) \to \\ & \quad \quad \exists x . \neg (\exists z_2 . (z_2 = x \wedge z_2 \in P \wedge y \in \ceil*{z_2})) \\
{\text{iff}}& \vdash \exists z_1 . (z_1 \not \in P \wedge \top) \to \tag{Proposition~\ref{prop:var1}, \ref{prop:ereplacement}, and Corollary~\ref{cor:top}}\\
& \quad \quad \exists x . \neg (\exists z_2 . (z_2 = x \wedge z_2 \in P \wedge \top)) \\
{\text{iff}}& \vdash \exists z_1 . (z_1 \not \in P) \to \exists x . \neg (\exists z_2 . (z_2 = x \wedge z_2 \in P)) \\
{\text{iff}}& \vdash \forall x . (\exists z_2 . (z_2 = x \wedge z_2 \in P)) \to \forall z_1 . (z_1 \in P) \\
{\text{if}}& \vdash \forall z_1 . (\forall x . (\exists z_2 . (z_2 = x \wedge z_2 \in P)) \to (z_1 \in P)) \\
{\text{if}}& \vdash \forall x . (\exists z_2 . (z_2 = x \wedge z_2 \in P)) \to (z_1 \in P).
\end{align*}
Since $\vdash \forall x . (\exists z_2 . (z_2 = x \wedge z_2 \in P)) \to \exists z_2 . (z_2 = z_1 \wedge z_2 \in P) $, it suffices to show
\begin{align*}
&\vdash \exists z_2 . (z_2 = z_1 \wedge z_2 \in P) \to (z_1 \in P) \\
{\text{iff}}& \vdash z_1 \not \in P \to \forall z_2 . (z_2 \neq z_1 \vee z_2 \not \in P) \\
{\text{if}}& \vdash \forall z_2 . (z_1 \not \in P \to z_2 \neq z_1 \vee z_2 \not \in P) \\
{\text{if}}& \vdash z_1 \not \in P \to z_2 \neq z_1 \vee z_2 \not \in P \\
{\text{if}}& \vdash z_2 = z_1 \wedge z_2 \in P \to z_1 \in P.
\end{align*}
And we proved~\eqref{eq:prop11-1}.

Similarly, to show~\eqref{eq:prop11-2}, 
\begin{align*}
&\vdash \floor*{P} \to \forall x . (x \in P)  \\
{\text{iff}}& \vdash \exists x . \neg \ceil*{x \wedge P} \to \ceil*{\neg P} \\
{\text{iff}}& \vdash \forall y . (y \in \exists x . \neg \ceil*{x \wedge P} \to \ceil*{\neg P}) \\
{\text{if}}& \vdash y \in \exists x . \neg \ceil*{x \wedge P} \to \ceil*{\neg P} \\
{\text{iff}}& \vdash \exists x . \neg \exists z_2 . (z_2 = x \wedge z_2 \in P) \to \exists z_1 . (z_1 \not \in P) \\
{\text{iff}}& \vdash \forall z_1 . (z_1 \in P) \to \exists z_2 . (z_2 = x \wedge z_2 \in P)\\
{\text{iff}}& \vdash x \in P \to \exists z_2 . (z_2 = x \wedge z_2 \in P).
\end{align*}
We proved~\eqref{eq:prop11-2}.

\paragraph{Remark} If $x$ occurs free in $P$, the result does not hold. For example, let $P$ be $upto(x)$ where $upto(\cdot)$ is interpreted to $upto(n) = \{0, 1, \dots, n\}$ on $\mathbb{N}$. 
\end{proof}

\paragraph{Remark} From Membership Introduction and Elimination inference rules and Proposition~\ref{prop:floor}, $\vdash P$ iff $\vdash \floor*{P}$.

\begin{prop}[Classification Reasoning] \label{prop:classificationreasoning}
	For any $P$ and $Q$, from $\vdash P \to Q$ and $\vdash \neg P \to Q$ deduce $\vdash Q$.
\end{prop}
\begin{proof}
	From $\vdash \neg P \to Q$ deduce $\vdash \neg Q \to P$. Notice that $\vdash P \to Q$, so we have $\vdash \neg Q \to Q$, i.e., $\vdash \neg \neg Q \vee Q$ which concludes the proof.
\end{proof}
\begin{cor} \label{cor:classificationreasoning}
	For any $P_1$, $P_2$, and $Q$ are patterns with $\vdash P_1 \vee P_2$, from $\vdash P_1 \to Q$ and $\vdash P_2 \to Q$, deduce $\vdash Q$.
\end{cor}

\begin{defn}[Predicate Pattern]
A pattern $P$ is called a predicate pattern or a predicate if $\vdash (P = \top) \vee (P = \bot)$.
\end{defn}
\paragraph{Remark} Predicate patterns are closed under all logic connectives.
\paragraph{Remark} For any $P$, $\ceil*{P}$ is a predicate pattern.

\begin{prop} \label{prop:017}
	$\vdash (\ceil*{P} = \bot) = (P = \bot)$ and $\vdash (\floor*{P} = \top) = (P = \top)$.
\end{prop}
\begin{proof}
	It is easy to prove one derivation from the other, so we only prove the first one. By Proposition~\ref{prop:eestablish}, it suffices to prove
	\begin{equation} \label{eq:prop:017:a}
	\vdash (\ceil*{P} = \bot) \to (P = \bot)
	\end{equation}
	and
	\begin{equation} \label{eq:prop:017:b}
	\vdash (P = \bot) \to (\ceil*{P} = \bot)
	\end{equation}
	
	The proof of~\eqref{eq:prop:017:b} is trivial and we left it as an exercise. We now prove~\eqref{eq:prop:017:a} through the following backward reasoning.
	
	\begin{align}
	&\vdash (\ceil*{P} = \bot) \to (P = \bot)  \nonumber \\
	\text{iff} \quad &\vdash \forall y . (y \in ((\ceil*{P} = \bot) \to (P = \bot))) \nonumber \\
	\text{if} \quad &\vdash y \in ((\ceil*{P} = \bot) \to (P = \bot)) \nonumber \\
	\text{iff} \quad &\vdash (y \in (\ceil*{P} = \bot) \to (y \in (P = \bot)). \label{eq:prop:017:a:1}
	\end{align}
	While for any pattern $Q$, 
	\begin{align*}
	&\vdash y \in (Q = \bot) \\
	\text{iff} \quad &\vdash y \in \neg \ceil*{\neg (Q \leftrightarrow \bot)} \\
	\text{iff} \quad &\vdash y \in \neg \ceil*{Q}\\
	\text{iff} \quad &\vdash \neg \exists z . (z \in Q \wedge y \in \ceil*{z}) \\
	\text{iff} \quad &\vdash \neg \exists z . (z \in Q) 
	\end{align*}
	So we continue to prove~\eqref{eq:prop:017:a:1} by showing
	\begin{align*}
&\vdash (y \in (\ceil*{P} = \bot)) \to (y \in (P = \bot))  \\
\text{iff} \quad &\vdash \neg \exists z . (z \in \ceil*{P}) \to \neg \exists z . (z \in P) \\
\text{iff} \quad &\vdash \exists z . (z \in P) \to  \exists z . (z \in \ceil*{P}) \\
\text{iff} \quad &\vdash \exists z . (z \in P) \to \exists z . (\exists z_1 . (z_1 \in P \wedge z \in \ceil*{z_1}))\\
\text{iff} \quad &\vdash \exists z . (z \in P) \to \exists z . \exists z_1 . (z_1 \in P) \\
\text{iff} \quad &\vdash \exists z_1 . (z_1 \in P) \to \exists z . \exists z_1 . (z_1 \in P).
	\end{align*}
	And we finish the proof by noticing the fact that for any pattern $Q$ and variable $x$, 
	$$ \vdash Q \to \exists x . Q.$$
	
\end{proof}



\begin{prop} \label{prop:predicateabsorb}
For any predicate $P$, $\vdash (P \neq \top) = (P = \bot)$ and $\vdash (P \neq \bot) = (P = \top)$.
\end{prop}
\begin{proof}
We only prove the first derivation, by showing both
\begin{equation}\label{eq:predicateabsort1}
\vdash (P \neq \top) \to (P = \bot)
\end{equation}
and
\begin{equation}\label{eq:predicateabsort2}
\vdash (P = \bot) \to (P \neq \top).
\end{equation}
Proving~\eqref{eq:predicateabsort2} is trivial. We now prove~\eqref{eq:predicateabsort1}, which is also trivial by transforming disjunction to implication.
\end{proof}
\begin{prop} \label{prop:predicatevee}
For any pattern $Q$ and any predicate pattern $P$, $\vdash P \vee Q$ iff $\vdash P \vee \floor*{Q}$.
\end{prop}
\begin{proof}
($\Leftarrow$) is obtained immediately by the remark of Proposition~\ref{prop:p->ceilp}. We now prove ($\Rightarrow$). 

Because $\vdash Q = \top \vee Q \neq \top$, it suffices to show
\begin{equation} \label{eq:predicatevee-01}
\vdash Q = \top \to (P \vee \floor*{Q} = \top)
\end{equation}
and
\begin{equation} \label{eq:predicatevee-02}
\vdash Q \neq \top \to (P \vee \floor*{Q} = \top)
\end{equation}
by Corollary~\ref{cor:classificationreasoning}, and the fact that $\vdash P \vee \floor*{Q} = \top$ and $\vdash \top$ imply $\vdash P \vee \floor*{Q}$. 

The proof of~\eqref{eq:predicatevee-01} is straightforward as follows.
\begin{align*}
&\vdash Q = \top \to (P \vee \floor*{Q} = \top) \\
\text{if} \quad &\vdash Q = \top \to (P \vee \floor*{\top} = \top) \\
\text{if} \quad &\vdash Q = \top \to (\top = \top) \\
\text{if} \quad &\vdash \top.
\end{align*}

The proof of~\eqref{eq:predicatevee-02} needs more effort:
\begin{align*}
&\vdash Q \neq \top \to (P \vee \floor*{Q} = \top) \\
\text{iff} \quad &\vdash (Q = \top) \vee (P \vee \floor*{Q} = \top) \\
\text{iff} \quad &\vdash (\floor*{Q} = \top) \vee (P \vee \floor*{Q} = \top) \\
\text{iff} \quad &\vdash \floor*{Q} \neq \top \to (P \vee \floor*{Q} = \top) \\
\text{iff} \quad &\vdash \floor*{Q} = \bot \to (P \vee \floor*{Q} = \top) \\
\text{if} \quad &\vdash \floor*{Q} = \bot \to (P \vee \bot = \top) \\
\text{iff} \quad &\vdash \floor*{Q} = \bot \to (P = \top) \\
\text{if} \quad &\vdash Q = \top \vee P = \top.
\end{align*}
Notice that $P$ is a predicate pattern, so it suffices to show
\begin{equation*}
\vdash P = \top \to (Q = \top \vee P = \top),
\end{equation*}
whose validity is obvious, and
\begin{equation*}
\vdash P = \bot \to (Q = \top \vee P = \top),
\end{equation*}
which is proved by showing
\begin{equation} 
\vdash P = \bot \to Q = \top.
\end{equation}
Because $\vdash P \vee Q$, it suffices to show
\begin{align*}
&\vdash P = \bot \to (P \vee Q) \to (Q = \top) \\
\text{if} \quad &\vdash P = \bot \to (\bot \vee Q) \to (Q = \top) \\
\text{iff} \quad &\vdash P = \bot \to Q \to (Q = \top) \\
\text{if} \quad &\vdash Q \to (Q = \top) \\
\text{iff} \quad &\vdash (Q \neq \top) \to \neg Q \\
\text{iff} \quad &\vdash (\floor*{Q} = \bot) \to \neg Q.
\end{align*}
Notice we have $\vdash Q \to \floor*{Q}$, which means $\vdash \neg \floor*{Q} \to \neg Q$, so it suffices to show
\begin{align*}
&\vdash (\floor*{Q} = \bot) \to \neg \floor*{Q} \\
\text{iff} \quad &\vdash (\floor*{Q} = \bot) \to \neg \bot \\
\text{iff} \quad &\vdash (\floor*{Q} = \bot) \to \top \\
\text{iff} \quad &\vdash \top.
\end{align*}
And this concludes the proof.
\end{proof}


\begin{prop}[Deduction Theorem]
If $\Gamma \cup \{P\} \vdash Q$ and the derivation does not use \mbox{$\forall x$-Generalization} where $x$ is free in $P$, then $\Gamma \vdash \floor*{P} \to Q$. 
\end{prop}
\begin{proof}
The proof is by induction on $n$, the length of the derivation of $Q$ from $\Gamma \cup \{P\}$.

Base step: $n = 1$, and $Q$ is an axiom, or $P$, or a member of $\Gamma$. If $Q$ is an axiom or a member of $\Gamma$, then $\Gamma \vdash Q$ and as a result, $\Gamma \vdash \floor*{P} \to Q$. If $Q$ is $P$, then $\Gamma \vdash \floor*{P} \to Q$ by Proposition~\ref{prop:p->ceilp}.

Induction step: Let $n > 1$. Suppose that if $P'$ can be deduced from $\Gamma \cup \{P\}$ without using $\forall x$-Generalization where $x$ is free in $P$, in a derivation containing fewer than $n$ steps, then $\Gamma \vdash \floor*{P} \to P'$.

Case 1: $Q$ is an axiom, or $P$, or a member of $\Gamma$. Precisely as in the Base step, we show that $\vdash \floor*{P} \to Q$.

Case 2: $Q$ follows from two previous patterns in the derivation by an application of Modus Ponens. These two patterns must have the forms $Q_1$ and $Q_1 \to Q$, and each one can certainly be deduced from $\Gamma \cup \{P\}$ by a derivation with fewer than $n$ steps, by just omitting the subsequent members from the original derivation from $\Gamma \cup \{P\} \vdash Q$. So we have $\Gamma \cup \{P\} \vdash Q_1$ and $\Gamma \cup \{P\} \vdash Q_1 \to Q$, and, applying the hypothesis of induction, $\Gamma \vdash \floor*{P} \to Q_1$ and $\Gamma \vdash \floor*{P} \to (Q_1 \to Q)$. It follows immediately that $\Gamma \vdash \floor*{P} \to Q$.

Case 3: $Q$ follows from a previous pattern in the derivation by an application of $\forall x_i$-Generalization where $x_i$ does not occur free in $P$. So $Q$ is $\forall x_i . Q_1$, say, and $Q_1$ appears previously in the derivation. Thus $\Gamma \cup \{P\} \vdash Q_1$, and the derivation has fewer than $n$ steps, so $\Gamma \vdash \floor*{P} \to Q_1$, since there is no application of Universal Generalization involving a free variable of $P$. Also $x_i$ cannot occur free in $P$, as it is involved in an application of Universal Generalization in the deduction of $Q$ from $\Gamma \cup \{P\}$. So we have a derivation of $\Gamma \vdash \floor*{P} \to Q$ as follows.
\begin{align*}
\Gamma &\vdash \floor*{P} \to Q \\
{\text{iff} \quad} \Gamma &\vdash \floor*{P} \to \forall x_i . Q_1 \\
{\text{if} \quad} \Gamma &\vdash \forall x_i . (\floor*{P} \to Q_1) \\
{\text{if} \quad} \Gamma &\vdash \floor*{P} \to Q_1. \\
\end{align*}
So $\Gamma \vdash \floor*{P} \to Q$ as required.

Case 4: Q follows from a previous pattern in the derivation by an application of Membership Introduction. So $Q$ is $\forall x_i . (x_i \in Q_1)$ with $x_i$ is free in $Q_1$, say, and $Q_1$ appears previously in the derivation. Thus $\Gamma \cup \{P\} \vdash Q_1$, and the derivation has fewer than $n$ steps, so $\Gamma \vdash \floor*{P} \to Q_1$, since there is no application of Universal Generalization involving a free variable of $P$. So we have a derivation of $\Gamma \vdash \floor*{P} \to Q$ as follows.
\begin{align*}
\Gamma &\vdash \floor*{P} \to Q \\
{\text{iff} \quad} \Gamma &\vdash \floor*{P} \to \forall x_i . (x_i \in Q_1) \\
{\text{iff} \quad} \Gamma &\vdash \floor*{P} \to \floor*{Q_1}, \\
\end{align*}
which follows by the hypothesis of induction $\Gamma \vdash \floor*{P} \to Q_1$ and the fact that $\Gamma \vdash Q_1 \to \floor*{Q_1}$ (by the Remark in Proposition~\ref{prop:p->ceilp}).

Case 5: Q follows from a previous pattern in the derivation by an application of Membership Elimination. The previous pattern must have the form $\forall x_i . (x_i \in Q)$, and can be deduced from $\Gamma \cup \{P\}$ by a derivation with fewer than $n$ steps, by just omitting the subsequent members from the original derivation from $\Gamma \cup \{P\} \vdash Q$. So we have $\Gamma \cup \{P\} \vdash \forall x_i . (x_i \in Q)$, and, applying the hypothesis of induction, $\Gamma \vdash \floor*{P} \to \forall x_i . (x_i \in Q)$. So we have a derivation of $\Gamma \vdash \floor*{P} \to Q$ as follows.
\begin{align*}
\Gamma &\vdash \floor*{P} \to Q \\
{\text{iff} \quad} \Gamma &\vdash \neg \floor*{P} \vee Q \\
{\text{iff} \quad} \Gamma &\vdash \neg \floor*{P} \vee \floor*{Q} \tag{Proposition~\ref{prop:predicatevee}} \\
{\text{iff} \quad} \Gamma &\vdash \neg \floor*{P} \vee \forall x_i . (x_i \in Q) \\
{\text{iff} \quad} \Gamma &\vdash \floor*{P} \to \forall x_i . (x_i \in Q),
\end{align*}
which is the hypothesis of induction. And this concludes our inductive proof.
\end{proof}

\begin{cor}[Closed-form Deduction Theorem]
If $P$ is closed, $\Gamma \cup \{P\} \vdash Q$ implies $\Gamma \vdash \floor*{P} \to Q$.
\end{cor}

\begin{thm} [Frame Rule]
	Let $\sigma \in \Sigma$ be a symbol in the signature. From $P_1 \to P_2$, deduce $\sigma(P_1) \to \sigma(P_2)$. In its most general form, $P_1 \to P_2$ deduces $\sigma(Q_1,\dots,P_1,\dots,Q_n) \to \sigma(Q_1,\dots, P_2, \dots, Q_n)$.
\end{thm}
\begin{proof}
	we write $\sigma(Q_1,\dots,P_i,\dots,Q_n)$ as $\sigma(P_i,\vec{Q})$ for short, for any $i \in \{1,2\}$.
	\begin{align*}
	&\vdash \sigma(P_1, \vec{Q}) \to \sigma(P_2,\vec{Q}) \\
	\text{iff} \quad &\vdash y \in (\sigma(P_1, \vec{Q}) \to \sigma(P_2,\vec{Q})) \\
	\text{iff} \quad &\vdash (y \in \sigma(P_1, \vec{Q})) \to (y \in \sigma(P_2, \vec{Q})) \\
	\text{iff} \quad &\vdash \exists z_1 . \exists \vec{z} . (z_1 \in P_1 \wedge \vec{z} \in \vec{Q} \wedge y \in \sigma(z_1, \vec{z})) \\
	&\to \exists z_2 . \exists \vec{z} . (z_2 \in P_2 \wedge \vec{z} \in \vec{Q} \wedge y \in \sigma(z_2, \vec{z})) \\
	\text{iff} \quad &\vdash \exists z_1 . \exists \vec{z} . (z_1 \in P_1 \wedge \vec{z} \in \vec{Q} \wedge y \in \sigma(z_1, \vec{z}) \\
	 &\quad \quad \quad \to z_1 \in P_2 \wedge \vec{z} \in \vec{Q} \wedge y \in \sigma(z_1, \vec{z})) \\
	\text{iff} \quad &\vdash \exists z_1. \exists \vec{z} . (z_1 \in P_1 \to z_1 \in P_2) \\
	\text{if} \quad &\vdash \exists z_1 .  (z_1 \in P_1 \to z_1 \in P_2) \\
	\text{if} \quad &\vdash P_1 \to P_2.
	\end{align*}
\end{proof}
\begin{cor} [Frame Rule as Implication]
	$\vdash \floor*{P \to Q} \to (\sigma(P) \to \sigma(Q))$
\end{cor}

\section{Inference rules}

\paragraph{Axioms}
$$
\infer{\Gamma \vdash A}{\cdot}
$$
where $A$ is an axiom. 

\paragraph{Inclusion}
$$
\infer{\Gamma \vdash P}{\cdot}
$$
where $P \in \Gamma$.

\paragraph{Modus Ponens}
$$
\infer{\Gamma \vdash P}
      {\Gamma \vdash Q \to P
      &\Gamma \vdash Q}
$$

\paragraph{Closed-Form Deduction Theorem}
$$
\infer{\Gamma \vdash P \to Q}
{\Gamma \cup \{P\} \vdash Q}
$$
where $P$ is closed.

\paragraph{Universal Generalization}
$$
\infer[(\forall x)]{\Gamma \vdash \forall x . P}
      {\Gamma \vdash P}
$$

\paragraph{Conjunction Splitting}
$$
\infer{\Gamma \vdash P \wedge Q}
      {\Gamma \vdash P
      &\Gamma \vdash Q}
$$

\end{document}
